---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readxl)
library(ggpubr)
library(rstatix)
library(dplyr)
library(forcats)
library(ggsci)
library(tibble)
library("phyloseq")
```

#Import tables
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip")

#read in otu table (smartchip normalized abundance)
otu_table <- read_excel("smartchip_phyloseq_tables.xlsx", sheet = "abundance")

#converting the abundance table to matrix (required to merge with taxonomy table)
otu_table <- otu_table %>% remove_rownames %>% column_to_rownames(var="GENE_ID") 
otu_table=as.matrix(otu_table,rownames=TRUE)

#multiplying abundance by 10^9 so we can calculate diversity measures (it won't be possible with decimals)
otu_table <- otu_table*10^9
mode(otu_table) <- "integer"

#converting empty values to 0 (in case they exist)
otu_table[is.na(otu_table)] <- 0

#read in taxonomy (gene ID for smartchip)
taxonomy <- read_excel("smartchip_phyloseq_tables.xlsx", sheet = "gene_ID")

#setting taxonomy row_names for GENE_ID (that will enable merging with the abundance table)
taxonomy <- taxonomy %>% remove_rownames %>% column_to_rownames(var="GENE_ID")
taxonomy=as.matrix(taxonomy)

#read in metadata
metadata <- read_excel("smartchip_phyloseq_tables.xlsx", sheet = "metadata")

#setting metadata row_names for sample ID (that will enable merging with the other tables)
metadata <- metadata %>% remove_rownames %>% column_to_rownames(var="Sample_ID")

#Changing the levels of source and group to make them better organized in the plots
metadata <- metadata %>% mutate(Source = fct_relevel(Source,"Meconium","Infant 7 days", "Mother"))
metadata <- metadata %>% mutate(Group = fct_relevel(Group,"Control","Sepsis"))

#import all tables as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)

# Same sample names
sample_names(OTU)
sample_names(META)

#Making phyloseq object
physeq=phyloseq(OTU,TAX,META)
```

#Saving smartchip as a dataframe format (metadata and abundance combined)
This serves for future analyses
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip")
physeq_df <- psmelt(physeq)
write.csv(physeq_df, "smartchip_data_frame.csv")
```

#Alpha diversity
We will calculate the Shannon and Observed genes
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq, measures = c("Shannon", "Observed"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq))
df <- reshape2::melt(df_alpha, measure.var=c("Shannon","Observed"),id.vars=c("Source","Group"))
df$value = as.numeric(df$value)
df <- df %>% mutate(Source = fct_relevel(Source,"Meconium","Infant 7 days", "Mother"))
df <- df %>% mutate(Group = fct_relevel(Group,"Control","Sepsis"))

#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df, variable == "Shannon")
observed = dplyr::filter(df, variable == "Observed")
```

##SHANNON
###Stats
Kruskal Wallis between groups
```{r}
kw.test <- shannon %>% 
  kruskal_test(value ~ Group)
kw.test
```
Wilcoxon test stratified by source between groups
```{r}
stat.test <- shannon %>% 
  group_by(Source) %>%
  wilcox_test(value ~ Group, p.adjust.method = "none", alternative = "less") %>%
  add_significance("p") %>%
  add_xy_position(x = "Source", dodge = 0.8)
stat.test
```

###Boxplot
```{r}
shannon_arg_boxplot = ggboxplot(shannon, x = "Source", y = "value", color = "Group", palette = "jama", fill = "Group", add = c("jitter"), notch = F, outlier.shape = NA) +labs(x = element_blank(), y = "Shannon index") + scale_fill_jama(alpha = 0.5) + theme(legend.position="top") +
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) + 
  stat_compare_means(comparisons = list(
    c("Meconium", "Infant 7 days"), 
    c("Infant 7 days", "Mother"), 
    c("Meconium", "Mother")
    ), 
    label.y = c(4,4.4,4.8), tip.length=.02, hide.ns = F, color = "gray35") +
    annotate("text", x = 2, y = 5.8, label = "Kruskal-Wallis, p = 0.154", colour = "black")
shannon_arg_boxplot
```
###Boxplot groups
```{r}
stat.test <- shannon %>% 
  group_by(Group) %>%
  wilcox_test(value ~ Source, p.adjust.method = "none", alternative = "less") %>%
  add_significance("p") %>%
  add_xy_position(x = "Source", dodge = 0.8)
stat.test
```

```{r}
shannon_arg_boxplot_groups = ggboxplot(shannon, x = "Source", y = "value", color = "Group", palette = "jama", fill = "Group", add = c("jitter"), notch = F, outlier.shape = NA) +
  labs(x = element_blank(), y = "Shannon index") + scale_fill_jama(alpha = 0.5) + 
  theme(legend.position="top") + 
  facet_wrap(vars(Group)) + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) 
shannon_arg_boxplot_groups
```

##observed
###Stats
Kruskal Wallis between groups
```{r}
kw.test <- observed %>% 
  kruskal_test(value ~ Group)
kw.test
```

Wilcoxon stratified by souce and compared between groups
```{r}
stat.test <- observed %>% 
  group_by(Source) %>%
  wilcox_test(value ~ Group, p.adjust.method = "none", alternative = "less") %>%
  add_significance("p") %>%
  add_xy_position(x = "Source", dodge = 0.8)
stat.test
```

###Boxplot
```{r}
observed_arg_boxplot = ggboxplot(observed, x = "Source", y = "value", color = "Group", palette = "jama", fill = "Group", add = c("jitter"), notch = F, outlier.shape = NA, alpha = 0.5) +labs(x = element_blank(), y = "Observed") + 
  theme(legend.position="top") + 
  stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) +
  stat_compare_means(comparisons = list(
    c("Meconium", "Infant 7 days"), 
    c("Infant 7 days", "Mother"), 
    c("Meconium", "Mother")
    ), 
    label.y = c(170,190,210), tip.length=.02, hide.ns = F, color = "gray35") +
    annotate("text", x = 2, y = 250, label = "Kruskal-Wallis, p = 0.654", colour = "black")

observed_arg_boxplot
```
###Boxplot groups
```{r}
stat.test <- observed %>% 
  group_by(Group) %>%
  wilcox_test(value ~ Source, p.adjust.method = "none", alternative = "less") %>%
  add_significance("p")
stat.test
```

```{r}
observed_arg_boxplot_groups = ggboxplot(observed, x = "Source", y = "value", color = "Group", palette = "jama", fill = "Group", add = c("jitter"), notch = F, outlier.shape = NA) +labs(x = element_blank(), y = "Nº Observed ARGs") + scale_fill_jama(alpha = 0.5) + theme(legend.position="top") + facet_wrap(vars(Group)) 

# Add p-values onto the box plots
stat.test <- stat.test %>%
  add_xy_position(x = "Source", dodge = 0.8)

observed_arg_boxplot_groups = observed_arg_boxplot_groups + stat_pvalue_manual(stat.test,  label = "p", tip.length = 0) 
observed_arg_boxplot_groups
```

#Figure Alpha-diversity ARGs
```{r fig.height=5, fig.width=8}
alpha_plots <-  ggarrange(shannon_arg_boxplot, observed_arg_boxplot, ncol = 2, labels = c("A", "B"), common.legend = T)
alpha_plots
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = alpha_plots, "alpha_diversity_smartchip.png",width=10,height=5)
```


##Beta-diversity
###Bray statistics
```{r}
## PERMANOVA
library(vegan)
bray_dist = phyloseq::distance(physeq, method="bray") #RUN this only once because it takes a lot of time
adonis_bray = adonis(bray_dist ~ sample_data(physeq)$Group_Source)
adonis_bray

adonis_bray = adonis(bray_dist ~ sample_data(physeq)$Source)
adonis_bray

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
bray_disp <-betadisper(bray_dist, sample_data(physeq)$Group_Source, type=c("median"))
anova(bray_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(bray_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(bray_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(bray_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(bray_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(bray_disp)

bray_disp_source <-betadisper(bray_dist, sample_data(physeq)$Source, type=c("median"))
anova(bray_disp_source)
TukeyHSD(bray_disp_source)

scores(bray_disp, 1:4, display = "centroids")

```
##Bray beta diversity with phyloseq
```{r}
beta_bray <- ordinate(physeq, "PCoA", "bray") #RUN this only ONCE because it takes a lot of time
```

##Bray beta diversity plot with phyloseq
Source
```{r}
beta_bray_plot_source = plot_ordination(physeq, beta_bray, type="Source", color="Source", shape="Source") +
  stat_ellipse(type = "t", lineSource = 3) + 
  theme_classic()+
  scale_fill_aaas() +
  scale_color_aaas() + 
  annotate("text", x = 0.3, y = 0.7, label = expression(paste("PERMANOVA, ",R^2 ,"= 0.07, ", paste(italic('p')),"=0.001"))) +
  theme(legend.position="top")
  #To add arrows https://neavemj.github.io/posts/coralMicrobiome +

beta_bray_plot_source
```
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = beta_bray_plot_source, "beta_bray_plot_source.png", width = 8, height = 5)
```

Group and source
```{r}
beta_bray_plot_groups = plot_ordination(physeq, beta_bray, type="Group_Source", color="Group_Source", shape="Group_Source") +
  stat_ellipse(type = "t", lineSource = 3) + 
  theme_classic()+
 # labs(colour = "Source") + 
  scale_fill_aaas() +
  scale_color_aaas() + 
  annotate("text", x = 0.3, y = 0.7, label = expression(paste("PERMANOVA, ",R^2 ,"= 0.07, ", paste(italic('p')),"=0.001"))) +
  theme(legend.position="top")
  #To add arrows https://neavemj.github.io/posts/coralMicrobiome +

beta_bray_plot_groups
```
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = beta_bray_plot_groups, "beta_bray_plot_group&source.png", width = 8, height = 5)
```

#COMPOSITION
##Class data frame
```{r}
physeq <- subset_taxa(physeq, Type%in%c("ARG"))
```

```{r}
glom_class <- tax_glom(physeq, taxrank = 'Class')
data_class <- psmelt(glom_class) # create dataframe from phyloseq object

library(dplyr)
data_class_rel = data_class %>% group_by(Sample) %>% mutate(relAbundBySample = Abundance / sum(Abundance)*100) #Converting counts to relative abundance %
data_class_rel = data_class_rel %>% mutate(Abundance = Abundance / 10^9) #Converting counts to normalized abundance since I multiplied the normalized abundance by 10^9 to calculate the Shannon indes
```

```{r fig.width=12}
ARG_class_ab <- ggbarplot(data=data_class_rel, x= "Sample", y = "relAbundBySample", color = "Class", fill = "Class", palette = get_palette("Paired",15), ylab = "Abundance", 
                          #order = metadata$Sample_ID
                          ) + 
  facet_wrap(~Source, scales = "free_x")+labs(colour = "Class", fill = "Class") + 
  theme(axis.text.x = element_text(angle = 90, size = 5, hjust = 1, colour = "gray35"))

ARG_class_ab
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = ARG_class_ab, "smartchip_class_relab_source.png", width = 20, height = 6)
```

```{r fig.width=12}
ARG_class_ab_group_facet <- data_class_rel %>% 
  mutate(across(Group_Source, factor, levels =c("Control_Meconium", "Sepsis_Meconium", "Control_Infant 7 days", "Sepsis_Infant 7 days", "Control_Mother", "Sepsis_Mother"))) %>%
  ggbarplot(x= "Sample", y = "relAbundBySample", color = "Class", fill = "Class", palette = get_palette("Paired",15), ylab = "Abundance") + 
  facet_wrap(~Group_Source, scales = "free_x", nrow = 3, ncol = 2)+
  labs(colour = "Class", fill = "Class") + 
  theme(axis.text.x = element_text(angle = 90, size = 5, hjust = 1, colour = "gray35"))

ARG_class_ab_group_facet
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = ARG_class_ab_group_facet, "smartchip_class_relab_group_source.png", width = 20, height = 10)
```

```{r}

ARG_class_ab_Group <- data_class_rel %>% 
  group_by(Group_Source, Class) %>% 
  summarise(Mean_abundance = mean(relAbundBySample)) %>% 
  ggbarplot(x= "Group_Source", y = "Mean_abundance", color = "Class", fill = "Class", palette = get_palette("Paired",15), 
            ylab = "Mean Relative Abundance", label.pos="out", position = position_stack(), xlab = "Group", orientation = "horiz", 
            order = c("Control_Meconium", "Sepsis_Meconium", "Control_Infant 7 days", "Sepsis_Infant 7 days", "Control_Mother", "Sepsis_Mother")) + 
  theme(legend.position = "right") +
  labs(colour = "Class", fill = "Class")
ARG_class_ab_Group
```
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/figures/smartchip")
ggsave(plot = ARG_class_ab_Group, "smartchip_class_mean_relab_group&source.png", width = 7, height = 4)
```


#ANCOM-BC
```{r}
library(microbiome)
library(nloptr)
library(ANCOMBC)
```


```{r}
physeq_meconium <-  subset_samples(physeq, Source %in% c("Meconium"))
physeq_7days <-  subset_samples(physeq, Source %in% c("Infant 7 days"))
physeq_mother <-  subset_samples(physeq, Source %in% c("Mother"))
```

##Class
```{r}
# Aggregate to phylum level
physeq = tax_glom(physeq, "Class")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa") 
  merge(day1_Phylum@tax_table,sigtab, by = 0)

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc/ANCOM_class_groups.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```
#Anaerobic Control vs. Ceftiofur
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_anero, "Class")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_class_anaerobic_con_cef.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Aerobic Control vs. Ceftiofur
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_aero, "Class")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_class_aerobic_con_cef.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

##Gene
#Initial Control vs. Ceftiofur
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_ini, "Gene")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_gene_ini_con_cef.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```
#Anaerobic Control vs. Ceftiofur
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_anero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_gene_anaerobic_con_cef.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Aerobic Control vs. Ceftiofur
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_aero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Group",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Group", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_gene_aerobic_con_cef.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```


##Class
#Control Initial vs Aerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_ini_aero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_class_ini_aerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Control Initial vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_ini_anero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_class_ini_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Control Aerobic vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_aero_anero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_control_class_aerobic_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Initial vs Aerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_ini_aero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_class_ini_aerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Initial vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_ini_anero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_class_ini_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Aerobic vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_aero_anero, "Class")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_class_aerobic_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Gene
##Control Initial vs Aerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_ini_aero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_control_gene_ini_aerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

##Control Initial vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_ini_anero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_control_gene_ini_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Control Aerobic vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_con_aero_anero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_control_gene_aerobic_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Initial vs Aerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_ini_aero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_gene_ini_aerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Initial vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_ini_anero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_gene_ini_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Ceftiofur Aerobic vs Anerobic
```{r}
# Aggregate to phylum level
physeq = aggregate_taxa(physeq_cef_aero_anero, "Gene")

out = ancombc(phyloseq = physeq, formula = "Source",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Source", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = TRUE)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)
ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- tibble::rownames_to_column(ancom_signif_names, "Taxa")

write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ANCOM_cef_gene_aerobic_anaerobic.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ")
#Tables to merge to add identifiers in gene
taxonomy_df <- read_excel("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Digesters_project/tables/smartchip/smartchip_phyloseq_tables.xlsx", sheet = "taxonomy")
ancom_tb <- read.csv("ANCOM_cef_gene_aerobic_anaerobic.csv") 

ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)

#Writing ANCOM table with identifiers
write.table(ancom_tb_id, file="ANCOM_cef_gene_aerobic_anaerobic.csv",sep = ",", row.names = F) 
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/ancombc ")

####Table 2 #####
ancom_tb <- read.csv("ANCOM_cef_gene_ini_aerobic.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_cef_gene_ini_aerobic.csv",sep = ",", row.names = F) 

####Table 3 #####
ancom_tb <- read.csv("ANCOM_cef_gene_ini_anaerobic.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_cef_gene_ini_anaerobic.csv",sep = ",", row.names = F) 

####Table 4 #####
ancom_tb <- read.csv("ANCOM_control_gene_aerobic_anaerobic.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_control_gene_aerobic_anaerobic.csv",sep = ",", row.names = F) 

####Table 5 #####
ancom_tb <- read.csv("ANCOM_control_gene_ini_aerobic.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_control_gene_ini_aerobic.csv",sep = ",", row.names = F) 

####Table 6 #####
ancom_tb <- read.csv("ANCOM_control_gene_ini_anaerobic.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_control_gene_ini_anaerobic.csv",sep = ",", row.names = F) 

####Table 7 #####
ancom_tb <- read.csv("ANCOM_gene_aerobic_con_cef.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_gene_aerobic_con_cef.csv",sep = ",", row.names = F) 

####Table 8 #####
ancom_tb <- read.csv("ANCOM_gene_anaerobic_con_cef.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_gene_anaerobic_con_cef.csv",sep = ",", row.names = F) 

####Table 9 #####
ancom_tb <- read.csv("ANCOM_gene_ini_con_cef.csv") 
ancom_tb_id <- merge(taxonomy_df,cef_ae_an,by="OTU_ID",all=F)
write.table(ancom_tb_id, file="ANCOM_gene_ini_con_cef.csv",sep = ",", row.names = F) 

```

#Alluvial plot
```{r fig.height=20, fig.width=10}
physeq_df <- psmelt(physeq_no_16s) # create dataframe from phyloseq object
physeq_df_rel_ab = physeq_df %>% group_by(Sample) %>% mutate(relAbundBySample = Abundance / sum(Abundance)*100) #Converting counts to relative abundance %



physeq_for_alluvia <- physeq_df_rel_ab %>% 
  select(Source,Class,Gene,relAbundBySample) %>% 
    filter(Source %notin% c("Taxanomic"))


library(ggalluvial)
alluvia_class <- ggplot(as.data.frame(physeq_for_alluvia),
       aes(y = relAbundBySample/18, axis1 = Source, axis2 = Class)) +
  geom_alluvium(aes(fill = Class), width = 1/12) +
  geom_stratum(width = 1/12, color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
#  scale_x_discrete(limits = c("Source", "Class","Gene"), expand = c(.05, .05,.05)) +
  scale_fill_simpsons() +
  ggtitle("Relative abundance of ARGs")
alluvia_class
```

```{r}
physeq_for_alluvia_ARGs <- physeq_df_rel_ab %>% 
  select(Source,Class,Gene,relAbundBySample) %>% 
  filter(Source %in% c("ARGs"))

physeq_for_alluvia_ARGs_no_tet <- physeq_for_alluvia_ARGs %>% 
  select(Source,Class,Gene,relAbundBySample) %>% 
  filter(Class %notin% c("Tetracycline"))

alluvia_class_ARGs <- ggplot(as.data.frame(physeq_for_alluvia_ARGs),
       aes(y = relAbundBySample/18, axis1 = Source, axis2 = Class, axis3 = Gene)) +
  geom_alluvium(aes(fill = Class), width = 1/12) +
  geom_stratum(width = 1/12, color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
#  scale_x_discrete(limits = c("Source", "Class","Gene"), expand = c(.05, .05,.05)) +
  scale_fill_brewer(Source = "qual", palette = "Set1") +
  ggtitle("Relative abundance of ARGs")
alluvia_class_ARGs

alluvia_class_ARGs_notet <- ggplot(as.data.frame(physeq_for_alluvia_ARGs_no_tet),
       aes(y = relAbundBySample/18, axis1 = Source, axis2 = Class, axis3 = Gene)) +
  geom_alluvium(aes(fill = Class), width = 1/12) +
  geom_stratum(width = 1/12, color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
#  scale_x_discrete(limits = c("Source", "Class","Gene"), expand = c(.05, .05,.05)) +
  scale_fill_brewer(Source = "qual", palette = "Set1") +
  ggtitle("Relative abundance of ARGs")
alluvia_class_ARGs_notet
```
#Heat-tree
#Modifying phyloseq object
```{r}
#To plot the tolal counts as percentage in the heat tree, divide the sum of the counts by the number of samples and multiply by 100. Here, I have 158 samples.

physeq_t  = transform_sample_counts(physeq_ARGs, function(x) x / sum(x)/18*100 ) 

#Filter all the taxa that have a percentage higher than 0.0001. This is optional, but I prefer to plot the most abundant genes/taxa

physeq_t1 = phyloseq::filter_taxa(physeq_t, function(x) mean(x) > 1e-5, TRUE) 
```

#Metacoder
```{r}
library(metacoder)

#transforming physeq object into a metacoder object

metacoder_object = parse_phyloseq(physeq_t1) 

#creating a column with total counts. Keep as is
metacoder_object$data$taxon_counts <- calc_taxon_abund(metacoder_object, data = "otu_table") 

metacoder_object$data$taxon_counts$total <- rowSums(metacoder_object$data$taxon_counts[, -1]) # -1 = taxon_id column

metacoder_object$data$taxon_prop <- calc_obs_props(metacoder_object, "otu_table")
```
#Heat tree
```{r}
#Plotting heat-tree

heat_tree = metacoder_object %>%
  # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = metacoder_object$data$taxon_counts$total,
             node_size_range = c(0.01, 0.06),
            edge_size_range = c(0.005, 0.005),
            node_label_size_range = c(0.01, 0.05),
         node_color = metacoder_object$data$taxon_counts$total,

            node_color_axis_label = "Proportion",
            node_size_axis_label = "ARG count",
            overlap_avoidance =1) #Modify if necssary
heat_tree
```

#PDF
```{r}
#Saving plot as PDF

#Change with your desired location
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Digesters_project/figures_plots/smartchip_ct32") 

# mm to inch
setWidth = 11

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='ARGs_heatree.pdf',width=setWidth,height=9,pointsize=setFontSize)
heat_tree
dev.off()
```
#Modifying phyloseq object
```{r}
#To plot the tolal counts as percentage in the heat tree, divide the sum of the counts by the number of samples and multiply by 100. Here, I have 158 samples.

physeq_t  = transform_sample_counts(physeq_MGEs, function(x) x / sum(x)/18*100 ) 

#Filter all the taxa that have a percentage higher than 0.0001. This is optional, but I prefer to plot the most abundant genes/taxa

physeq_t1 = phyloseq::filter_taxa(physeq_t, function(x) mean(x) > 1e-5, TRUE) 
```

#Metacoder
```{r}
library(metacoder)

#transforming physeq object into a metacoder object

metacoder_object = parse_phyloseq(physeq_t1) 

#creating a column with total counts. Keep as is
metacoder_object$data$taxon_counts <- calc_taxon_abund(metacoder_object, data = "otu_table") 

metacoder_object$data$taxon_counts$total <- rowSums(metacoder_object$data$taxon_counts[, -1]) # -1 = taxon_id column

metacoder_object$data$taxon_prop <- calc_obs_props(metacoder_object, "otu_table")
```
#Heat tree
```{r}
#Plotting heat-tree

heat_tree = metacoder_object %>%
  # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = metacoder_object$data$taxon_counts$total,
             node_size_range = c(0.01, 0.06),
            edge_size_range = c(0.005, 0.005),
            node_label_size_range = c(0.01, 0.05),
         node_color = metacoder_object$data$taxon_counts$total,

            node_color_axis_label = "Proportion",
            node_size_axis_label = "MGE count",
            overlap_avoidance =1) #Modify if necssary
heat_tree
```

#PDF
```{r}
#Saving plot as PDF

#Change with your desired location
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Digesters_project/figures_plots/smartchip_ct32") 

# mm to inch
setWidth = 11

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='MGEs_heatree.pdf',width=setWidth,height=9,pointsize=setFontSize)
heat_tree
dev.off()
```


