---
title: "R Notebook"
output: html_notebook
---


#Import tables
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip")

library(readxl)
#gene abundance
gene_ab <- read_excel("smartchip_maaslin2_tables.xlsx", sheet = "gene") %>% 
  remove_rownames %>% 
  column_to_rownames(var="Gene") 

#Class abundance
class_ab <- read_excel("smartchip_maaslin2_tables.xlsx", sheet = "class") %>% 
  remove_rownames %>% 
  column_to_rownames(var="Class")

#read in metadata
metadata <- read_excel("smartchip_maaslin2_tables.xlsx", sheet = "metadata") %>% 
  remove_rownames %>% 
  column_to_rownames(var="Sample_ID")
```

#MaASlin2
```{r}
#INSTALLATION https://github.com/biobakery/Maaslin2

if(!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("Maaslin2")
```

#Global comparison any antibiotic use
Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = gene_ab,
    input_metadata = metadata, 
    output = 'antibiotic_use_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = class_ab,
    input_metadata = metadata, 
    output = 'antibiotic_use_class', 
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Filtering meconium samples in metadata and abundance tables
```{r}
meconium_metadata <- metadata %>% filter(Source == "Meconium")
meconium_samples <- metadata %>% filter(Source == "Meconium") %>% rownames() %>% as.data.frame() %>% rename(sample = ".")
meconium_gene_ab <- gene_ab %>% select(meconium_samples$sample)
meconium_class_ab <- class_ab %>% select(meconium_samples$sample)
```

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = meconium_gene_ab,
    input_metadata = meconium_metadata, 
    output = 'meconium_antibiotic_use_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = meconium_class_ab,
    input_metadata = meconium_metadata, 
    output = 'meconium_antibiotic_use_class', 
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Filtering Infant 7 days samples in metadata and abundance tables
```{r}
infant7days_metadata <- metadata %>% filter(Source == "Infant 7 days")
infant7days_samples <- metadata %>% filter(Source == "Infant 7 days") %>% rownames() %>% as.data.frame() %>% rename(sample = ".")
infant7days_gene_ab <- gene_ab %>% select(infant7days_samples$sample)
infant7days_class_ab <- class_ab %>% select(infant7days_samples$sample)
```

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = infant7days_gene_ab,
    input_metadata = infant7days_metadata, 
    output = 'infant7days_antibiotic_use_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = infant7days_class_ab,
    input_metadata = infant7days_metadata, 
    output = 'infant7days_antibiotic_use_class', 
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Filtering Mother samples in metadata and abundance tables
```{r}
mother_metadata <- metadata %>% filter(Source == "Mother")
mother_samples <- metadata %>% filter(Source == "Mother") %>% rownames() %>% as.data.frame() %>% rename(sample = ".")
mother_gene_ab <- gene_ab %>% select(mother_samples$sample)
mother_class_ab <- class_ab %>% select(mother_samples$sample)
```

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = mother_gene_ab,
    input_metadata = mother_metadata, 
    output = 'mother_antibiotic_use_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = mother_class_ab,
    input_metadata = mother_metadata, 
    output = 'mother_antibiotic_use_class', 
    fixed_effects = c('Any_Ab_use'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Global comparison GROUPS (sepsis and control)
Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = gene_ab,
    input_metadata = metadata, 
    output = 'Group_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = class_ab,
    input_metadata = metadata, 
    output = 'Group_class', 
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

##Meconium

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = meconium_gene_ab,
    input_metadata = meconium_metadata, 
    output = 'meconium_Group_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = meconium_class_ab,
    input_metadata = meconium_metadata, 
    output = 'meconium_Group_class', 
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Filtering Infant 7 days samples in metadata and abundance tables
```{r}
infant7days_metadata <- metadata %>% filter(Source == "Infant 7 days")
infant7days_samples <- metadata %>% filter(Source == "Infant 7 days") %>% rownames() %>% as.data.frame() %>% rename(sample = ".")
infant7days_gene_ab <- gene_ab %>% select(infant7days_samples$sample)
infant7days_class_ab <- class_ab %>% select(infant7days_samples$sample)
```

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = infant7days_gene_ab,
    input_metadata = infant7days_metadata, 
    output = 'infant7days_Group_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = infant7days_class_ab,
    input_metadata = infant7days_metadata, 
    output = 'infant7days_Group_class', 
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

#Filtering Mother samples in metadata and abundance tables
```{r}
mother_metadata <- metadata %>% filter(Source == "Mother")
mother_samples <- metadata %>% filter(Source == "Mother") %>% rownames() %>% as.data.frame() %>% rename(sample = ".")
mother_gene_ab <- gene_ab %>% select(mother_samples$sample)
mother_class_ab <- class_ab %>% select(mother_samples$sample)
```

Gene level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = mother_gene_ab,
    input_metadata = mother_metadata, 
    output = 'mother_Group_gene', #Generates a directory with that name that includes all the results (figures and tables)
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```

Class level
```{r}
library(Maaslin2)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Mila_project/tables/smartchip/maaslin2/")
fit_data <- Maaslin2(
    input_data = mother_class_ab,
    input_metadata = mother_metadata, 
    output = 'mother_Group_class', 
    fixed_effects = c('Group'),
    random_effects = c("Deliver_mode"),
    standardize = TRUE)
```